---
title: "esm206_week9"
author: "Deanna Mireles"
date: "11/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(broom)
library(janitor)
```

## Read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts od lizards by species
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n) # Shows the list from most to least
```

## Make a subset

- UTST lizards
- Sites: IBPE (enclosure grassland site), RABB (rabbit mesquite site), "EAST" (east tarbush site)
- Tail condition recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

## Find counts and proportions of lizards with borken or whole tails at each site

```{r}
# First Way - long format doesn't look very nice
utst %>% 
  count(site, tail)
```

```{r}
# Second Way- easier and gives results in contingency table format; better for chi-squared function
tail_counts <- utst %>% 
  tabyl(site, tail) 


# Another version w/ proportion by row **
tail_proportions <- tail_counts %>% 
  adorn_percentages() %>% # Gives count by %
  adorn_pct_formatting(digits = 2) %>% # Changed sig fig to 2
  adorn_ns() # Actual counts included as well
```

## Convert to a contingency table

Use `column_to_rowname()`.

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Run Chi-square

Are tail breakage and site independent? (if yes, site has no impact on tail breakages and we will see similar proportions across all three sites)

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- tidy(lizard_x2)
```

X-squared = 13.064, p-value = 0.001456 -> p < 0.05 so, yes we have sufficient evident to reject he null hypothesis of independence site and tail breakage are not independent. There is a significant association between site and tail breakage.

Say it like this w/ inline reference: There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r round(lizard_tidy$statistic, 2)`, *p* = `r round(lizard_tidy$p.value, 2)`).


